---
title: "model fitting"
author: "shahaf"
date: "2024-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(janitor)
library(kableExtra)
library(dplyr)
library(car)
library(infer)
library(boot)
```

```{r, include=FALSE}
VacantApt <- read.csv("../data/Updated/vacant_apartments.csv")
PropertiesRooms <- read.csv("../data/Updated/properties_by_rooms.csv")
Purchases <- read.csv("../data/Updated/property_purchases.csv")

Kneset20 <- read.csv("../data/Updated/Kneset20_summed.csv")
Kneset21 <- read.csv("../data/Updated/Kneset21_summed.csv")
Kneset22 <- read.csv("../data/Updated/Kneset22_summed.csv")
Kneset23 <- read.csv("../data/Updated/Kneset23_summed.csv")
Kneset24 <- read.csv("../data/Updated/Kneset24_summed.csv")
Kneset25 <- read.csv("../data/Updated/Kneset25_summed.csv")
```


```{r, include=FALSE}
VacantApt
PropertiesRooms
Purchases
```


```{r, include=FALSE}
Kneset20
Kneset21
Kneset22
Kneset23
Kneset24
Kneset25
```


```{r, include=FALSE}
PropertiesRooms <- drop_na(PropertiesRooms) # 4 apartments who add no floor data was removed
# replacing NA's with 0's when there where no apartment purchased.
Purchases <- Purchases %>%
  mutate(
    X3.rooms.apartments = ifelse(is.na(X3.rooms.apartments), 0, X3.rooms.apartments),
    X4..rooms.apartments = ifelse(is.na(X4..rooms.apartments), 0, X4..rooms.apartments),
    average.price..NIS..3.rooms.apartments = ifelse(is.na(average.price..NIS..3.rooms.apartments), 0, average.price..NIS..3.rooms.apartments),
    average.price..NIS..4..rooms.apartments = ifelse(is.na(average.price..NIS..4..rooms.apartments), 0, average.price..NIS..4..rooms.apartments)
  )

Purchases <- Purchases %>% clean_names()
Purchases
```



```{r, include=FALSE}
# filter purchuses by year and join with relevant elections csv
Joined_Purch_Kneset20 <- Purchases %>%
  filter(year < 2019) %>%
  left_join(Kneset20, by = c("lamas_code" = "Lamas_code"))

Joined_Purch_Kneset21 <- Purchases %>%
  filter(year == 2019) %>%
  left_join(Kneset21, by = c("lamas_code" = "Lamas_code"))

Joined_Purch_Kneset22 <- Purchases %>%
  filter(year == 2020) %>%
  left_join(Kneset22, by = c("lamas_code" = "Lamas_code"))

Joined_Purch_Kneset23 <- Purchases %>%
  filter(year == 2021) %>%
  left_join(Kneset23, by = c("lamas_code" = "Lamas_code"))

Joined_Purch_Kneset24 <- Purchases %>%
  filter(year == 2022) %>%
  left_join(Kneset24, by = c("lamas_code" = "Lamas_code"))

Joined_Purch_Kneset25 <- Purchases %>%
  filter(year > 2022) %>%
  left_join(Kneset25, by = c("lamas_code" = "Lamas_code"))

# combining results
Purchuses_Elections <- bind_rows(Joined_Purch_Kneset20, Joined_Purch_Kneset21, Joined_Purch_Kneset22, Joined_Purch_Kneset23, Joined_Purch_Kneset24, Joined_Purch_Kneset25)
```


```{r}
Purchuses_Elections <- Purchuses_Elections %>%
  mutate(
    x3_average_paid_in_Mil = (x3_rooms_apartments * average_price_nis_3_rooms_apartments) / 1000000,
    x4_average_paid_in_Mil = (x4_rooms_apartments * average_price_nis_4_rooms_apartments) / 1000000,
    total_paid_in_Mil = x3_average_paid_in_Mil + x4_average_paid_in_Mil,
    sum_apartments = x3_rooms_apartments + x4_rooms_apartments
  )
```


```{r}
# selecting relevant columns
Purchuses_Elections <- Purchuses_Elections %>%
  dplyr::select(lamas_name, year, x3_rooms_apartments, x4_rooms_apartments, x3_average_paid_in_Mil, x4_average_paid_in_Mil, sum_apartments, total_paid_in_Mil, first.party, second.party, third.party, forth.party, first.party.percentage, second.party.percentage, third.party.percentage, forth.party.percentage, coalition.support, BZB, Vote.precentage, average_price_nis_3_rooms_apartments, average_price_nis_4_rooms_apartments)

# Fix party name
Purchuses_Elections <- Purchuses_Elections %>%
  mutate(
    first.party = case_when(
    first.party == "Kahul_lavan" ~ "Kahul_Lavan",
    first.party == "Yahadut_Hator" ~ "Yahadut_Hatora",
    first.party == "Israel_beitenu" ~ "Israel_Beitenu",
    first.party == "Zionot_datit" ~ "Tsionot_Datit",
    TRUE ~ first.party
    ),
    second.party = case_when(
    second.party == "Kahul_lavan" ~ "Kahul_Lavan",
    second.party == "Yahadut_Hator" ~ "Yahadut_Hatora",
    second.party == "Israel_beitenu" ~ "Israel_Beitenu",
    second.party == "Zionot_datit" ~ "Tsionot_Datit",
    TRUE ~ second.party
    ),
    third.party = case_when(
    third.party == "Kahul_lavan" ~ "Kahul_Lavan",
    third.party == "Yahadut_Hator" ~ "Yahadut_Hatora",
    third.party == "Israel_beitenu" ~ "Israel_Beitenu",
    third.party == "Zionot_datit" ~ "Tsionot_Datit",
    TRUE ~ third.party
    ),
    forth.party = case_when(
    forth.party == "Kahul_lavan" ~ "Kahul_Lavan",
    forth.party == "Yahadut_Hator" ~ "Yahadut_Hatora",
    forth.party == "Israel_beitenu" ~ "Israel_Beitenu",
    forth.party == "Zionot_datit" ~ "Tsionot_Datit",
    TRUE ~ forth.party
    )
  )

# apply factor
Purchuses_Elections <- Purchuses_Elections %>%
  mutate(
    first.party = as.factor(first.party),
    second.party = as.factor(second.party),
    third.party = as.factor(third.party),
    forth.party = as.factor(forth.party),
    )

```


```{r, include=FALSE}
PropertiesRooms <- PropertiesRooms %>%
  mutate(
    Total_Rooms = NoRooms + OneRooms + TwoRooms + ThreeRooms + FourRooms + FiveRooms + SixRooms + SevenRooms + EightRooms + NineRooms + TenRooms + MoreRooms
  )
PropertiesRooms
```


```{r}
# Group by lamas_code and sum Total_Rooms for each city in PropertiesRooms
PropertiesRooms_Summary <- PropertiesRooms %>%
  group_by(lamas_code, CityLMSName) %>%
  summarise(Total_Rooms_Sum = sum(Total_Rooms, na.rm = TRUE)) %>%
  ungroup()

# Join with Kneset25 and replace NA's in Total_Rooms with 0
Joined_Rooms_Kneset25 <- PropertiesRooms_Summary %>%
  mutate(lamas_code = as.numeric(lamas_code)) %>%
  right_join(Kneset25, by = c("lamas_code" = "Lamas_code")) %>%
  select(lamas_code, CityLMSName, Total_Rooms = Total_Rooms_Sum, first.party, first.party.percentage,
         second.party, second.party.percentage, third.party, third.party.percentage,
         forth.party, forth.party.percentage, coalition.support) %>%
  mutate(Total_Rooms = ifelse(is.na(Total_Rooms), 0, Total_Rooms))

# Print or use Joined_Rooms_Kneset25 as needed
Joined_Rooms_Kneset25

# # Prepare PropertiesRooms for model fitting
# Joined_Rooms_Kneset25 <- PropertiesRooms %>%
#   select(lamas_code, CityLMSName, Total_Rooms) %>%
#   mutate(lamas_code = as.numeric(lamas_code)) %>%
#   right_join(Kneset25, by = c("lamas_code" = "Lamas_code"))
# 
# Joined_Rooms_Kneset25 <- Joined_Rooms_Kneset25 %>%
#   select(lamas_code, CityLMSName, Total_Rooms, first.party, first.party.percentage, second.party, second.party.percentage, third.party, third.party.percentage, forth.party, forth.party.percentage, coalition.support) %>%
#   mutate(Total_Rooms = ifelse(is.na(Total_Rooms), 0, Total_Rooms))
# 
# Joined_Rooms_Kneset25
```


```{r, include=FALSE}
Joined_Vacant_Kneset25 <- VacantApt %>%
  select(CityLmsCode, CityLmsName) %>%
  group_by(CityLmsCode, CityLmsName) %>%
  summarise(Count_Vacant = n()) %>%
  left_join(Kneset25, by = c("CityLmsCode" = "Lamas_code"))

Joined_Vacant_Kneset25 <- Joined_Vacant_Kneset25 %>%
  select(CityLmsCode, CityLmsName, Count_Vacant, first.party, first.party.percentage, second.party, second.party.percentage, third.party, third.party.percentage, forth.party, forth.party.percentage, coalition.support)

Joined_Vacant_Kneset25
```


```{r}
# pearson correlation
Purchuses_Elections %>%
  dplyr::select(year, x3_rooms_apartments, x4_rooms_apartments, total_paid_in_Mil, coalition.support, average_price_nis_3_rooms_apartments, average_price_nis_4_rooms_apartments) %>%
  cor()
```


```{r, include=FALSE}
Purchuses_Elections
```


# Models - Purchases

```{r, include=FALSE}
# linear regression: predict total amount spent per city by the coalition support feature
coalition_totPaid_model <- lm(total_paid_in_Mil ~ coalition.support, data = Purchuses_Elections)
summary(coalition_totPaid_model)
plot(Purchuses_Elections$coalition.support, Purchuses_Elections$total_paid_in_Mil, 
     xlab = "Coalition Support", ylab = "Total Paid (in Million)", 
     main = "Linear Regression: Total Paid vs Coalition Support")
abline(coalition_totPaid_model, col = "blue")
```

```{r, include=FALSE}
# # bootstrapping the model
# lm_model <- function(data, indices) {
#   d <- data[indices, ] # allows boot to select sample
#   fit <- lm(total_paid_in_Mil ~ coalition.support, data = d)
#   return(coef(fit))
# }
# 
# set.seed(123) # for reproducibility
# boot_results <- boot(data = Purchuses_Elections, statistic = lm_model, R = 15000)
# boot_results
# 
# boot.ci(boot_results, type = "perc", index = 1) # Intercept
# boot.ci(boot_results, type = "perc", index = 2) # coalition.support
```

```{r, include=FALSE}
# linear regression: predict total amount spent per city by the top four party percentage
percentage_totPaid_model <- lm(total_paid_in_Mil ~ first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage + coalition.support, data = Purchuses_Elections)
summary(percentage_totPaid_model)
crPlots(percentage_totPaid_model)
```


```{r, include=FALSE}
# lm_model_percentage <- function(data, indices) {
#   d <- data[indices, ] # allows boot to select sample
#   fit <- lm(total_paid_in_Mil ~ first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage + coalition.support, data = d)
#   return(coef(fit))
# }
# 
# set.seed(123) # for reproducibility
# boot_results_percentage <- boot(data = Purchuses_Elections, statistic = lm_model_percentage, R = 15000)
# boot_results_percentage
# 
# ci_intercept <- boot.ci(boot_results_percentage, type = "perc", index = 1)
# ci_first_party <- boot.ci(boot_results_percentage, type = "perc", index = 2)
# ci_second_party <- boot.ci(boot_results_percentage, type = "perc", index = 3)
# ci_third_party <- boot.ci(boot_results_percentage, type = "perc", index = 4)
# ci_forth_party <- boot.ci(boot_results_percentage, type = "perc", index = 5)
# ci_coalition_support <- boot.ci(boot_results_percentage, type = "perc", index = 6)
# 
# # Print confidence intervals
# list(ci_intercept = ci_intercept, ci_first_party = ci_first_party, ci_second_party = ci_second_party, 
#      ci_third_party = ci_third_party, ci_forth_party = ci_forth_party, ci_coalition_support = ci_coalition_support)
```



```{r}
# poisson regression: predict log of num apartments purchased by coalition support
coalition_sumApt_model <- glm(sum_apartments ~ coalition.support, data = Purchuses_Elections, family = poisson)
summary(coalition_sumApt_model)
```

### log(sum_apartments) = 2.33383 + 0.34278 * coalition.support  
### This model is statistacly significant showing that the government buys more apartments if you support the coalition


```{r}
# poisson regression: predict log of num apartments purchased by coalition support and party percentage
coalition_partyPer_sumApt_model <- glm(sum_apartments ~ first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage + coalition.support, data = Purchuses_Elections, family = poisson)
summary(coalition_partyPer_sumApt_model)
```

### this model presents a better fit comparing to the previus model, we still see that higher coalition.support leads to more apartments purchases, but strong support of a party decreases that amount.


```{r}
# poisson regression: predict log of num apartments purchased by party percentage
partyPer_sumApt_model <- glm(sum_apartments ~ first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage, data = Purchuses_Elections, family = poisson)
summary(partyPer_sumApt_model)
```

### this model is poorer (by AIC) than the previous model, yet shows that strong support in top 3 parties decreases the amount of apartments purchases


```{r}
Party_sum_model <- glm(sum_apartments ~ first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage + coalition.support + first.party + second.party + third.party + forth.party, data = Purchuses_Elections, family = poisson)
summary(Party_sum_model)
```

### this model fits the data better than all previous models (by AIC and residual deviance), this model strethen previous results that stronger coalition support increases the amount of purchased apartmnets. Yet, strong support of one party decreases the amount of purchased apartments.


```{r}
# poisson regression: predict log of num apartments purchased by coalition support and vote percentage
coalition_votePer_model <- glm(sum_apartments ~ coalition.support + Vote.precentage, data = Purchuses_Elections, family = poisson)
summary(coalition_votePer_model)
```


# Models - PropertiesRooms

```{r, include=FALSE}
Joined_Rooms_Kneset25 <- Joined_Rooms_Kneset25 %>%
  mutate(ln_Total_Rooms = ifelse(Total_Rooms == 0, 0, log(Total_Rooms)))

Joined_Rooms_Kneset25
```

```{r, include=FALSE}
ln_TotRooms_Coalition_model <- lm(ln_Total_Rooms ~ coalition.support, data = Joined_Rooms_Kneset25)
summary(ln_TotRooms_Coalition_model)
```


```{r}
# linear regression: predict total amount of rooms per city by the coalition support feature
TotRooms_Coalition_model <- lm(Total_Rooms ~ coalition.support, data = Joined_Rooms_Kneset25)
summary(TotRooms_Coalition_model)
```

### Though the model is statisticaly significant, it has poor R-squared


```{r, include=FALSE}
# poisson regression: predict log of num apartments available by coalition support
coalition_totRooms_model <- glm(Total_Rooms ~ coalition.support, data = Joined_Rooms_Kneset25, family = poisson)
summary(coalition_totRooms_model)
```



```{r, include=FALSE}
# poisson regression: predict log of num apartments available by coalition support and party percentage
coalition_totRooms_partyPer_model <- glm(Total_Rooms ~ coalition.support + first.party.percentage + second.party.percentage + third.party.percentage + forth.party.percentage, family = poisson, data = Joined_Rooms_Kneset25)
summary(coalition_totRooms_partyPer_model)
```


# Alternative explenatory
```{r}
alternative_model1 <- lm(x3_rooms_apartments ~ average_price_nis_3_rooms_apartments, data = Purchuses_Elections)
summary(alternative_model1)

alternative_model2 <- lm(x4_rooms_apartments ~ average_price_nis_4_rooms_apartments, data = Purchuses_Elections)
summary(alternative_model2)

plot(Purchuses_Elections$average_price_nis_3_rooms_apartments, Purchuses_Elections$x3_rooms_apartments, 
     xlab = "average price", ylab = "3 rooms apartments", 
     main = "Linear Regression: 3 rooms apartments vs average price")
abline(alternative_model1, col = "blue")
abline(alternative_model2, col = "red")
```

### Significant but very poor R-squared, not a good explenatory variable.


# Model visualization
### chosen model: coalition_sumApt_model, df: Purchuses_Elections



```{r}
# Fit Poisson regression model
coalition_sumApt_model <- glm(sum_apartments ~ coalition.support, data = Purchuses_Elections, family = poisson)

# Predict values
predicted <- predict(coalition_sumApt_model, type = "response")

# Calculate errors
MAE <- mean(abs(Purchuses_Elections$sum_apartments - predicted))
RMSE <- sqrt(mean((Purchuses_Elections$sum_apartments - predicted)^2))

# Display results
cat("Mean Absolute Error (MAE):", MAE, "\n")
cat("Root Mean Squared Error (RMSE):", RMSE, "\n")
```


```{r}
# Poisson glm for a the data
ggplot(data = Purchuses_Elections, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(x = coalition.support, y = sum_apartments), se = TRUE, method = 'glm', method.args = list(family = 'poisson')) +
  labs(title = "Coalition Support vs Sum of Apartments",
       x = "Coalition Support",
       y = "Sum of Apartments",
    ) + 
  theme_minimal()
```

```{r}
# Poisson glm colored by year
ggplot(data = Purchuses_Elections, aes(x = coalition.support, y = sum_apartments, color = as.factor(year))) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(x = coalition.support, y = sum_apartments), se = FALSE, method = 'glm', method.args = list(family = 'poisson')) +
  labs(title = "Coalition Support vs Sum of Apartments",
       x = "Coalition Support",
       y = "Sum of Apartments",
       color = "Year") + 
  theme_minimal()
```

```{r}
# Poisson glm facet by year
ggplot(data = Purchuses_Elections, aes(x = coalition.support, y = sum_apartments, color = as.factor(year))) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(x = coalition.support, y = sum_apartments), se = TRUE, method = 'glm', method.args = list(family = 'poisson')) +
  labs(title = "Coalition Support vs Sum of Apartments",
       x = "Coalition Support",
       y = "Sum of Apartments",
       color = "Year") + 
  theme_minimal() +
  facet_wrap(~ year, ncol = 4) +
  theme(panel.spacing = unit(1, "lines"))
```

```{r}
# bar plot for 2017

# Filter data for 2017
data_year_2017 <- Purchuses_Elections %>% filter(year == 2017)

# Fit the model
model_2017 <- glm(sum_apartments ~ coalition.support, data = data_year_2017, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2017, type = "response")
rmse <- sqrt(mean((data_year_2017$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2017, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#CB485C") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#CB485C"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2018

# Filter data for 2018
data_year_2018 <- Purchuses_Elections %>% filter(year == 2018)

# Fit the model
model_2018 <- glm(sum_apartments ~ coalition.support, data = data_year_2018, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2018, type = "response")
rmse <- sqrt(mean((data_year_2018$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2018, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#CB0511") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#CB0511"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2019

# Filter data for 2019
data_year_2019 <- Purchuses_Elections %>% filter(year == 2019)

# Fit the model
model_2019 <- glm(sum_apartments ~ coalition.support, data = data_year_2019, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2019, type = "response")
rmse <- sqrt(mean((data_year_2019$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2019, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#B4010B") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#B4010B"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2020

# Filter data for 2020
data_year_2020 <- Purchuses_Elections %>% filter(year == 2020)

# Fit the model
model_2020 <- glm(sum_apartments ~ coalition.support, data = data_year_2020, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2020, type = "response")
rmse <- sqrt(mean((data_year_2020$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2020, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#740410") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#740410"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2021

# Filter data for 2021
data_year_2021 <- Purchuses_Elections %>% filter(year == 2021)

# Fit the model
model_2021 <- glm(sum_apartments ~ coalition.support, data = data_year_2021, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2021, type = "response")
rmse <- sqrt(mean((data_year_2021$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2021, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#CB485C") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#CB485C"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2022

# Filter data for 2022
data_year_2022 <- Purchuses_Elections %>% filter(year == 2022)

# Fit the model
model_2022 <- glm(sum_apartments ~ coalition.support, data = data_year_2022, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2022, type = "response")
rmse <- sqrt(mean((data_year_2022$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2022, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#CB0511") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#CB0511"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# bar plot for 2023

# Filter data for 2023
data_year_2023 <- Purchuses_Elections %>% filter(year == 2023)

# Fit the model
model_2023 <- glm(sum_apartments ~ coalition.support, data = data_year_2023, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2023, type = "response")
rmse <- sqrt(mean((data_year_2023$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2023, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#B4010B") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#B4010B"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

```{r}
# Filter data for 2024
data_year_2024 <- Purchuses_Elections %>% filter(year == 2024)

# Fit the model
model_2024 <- glm(sum_apartments ~ coalition.support, data = data_year_2024, family = poisson())

# Calculate the RMSE
predicted <- predict(model_2024, type = "response")
rmse <- sqrt(mean((data_year_2024$sum_apartments - predicted)^2))

# Create ggplot with enhanced aesthetics
ggplot(data_year_2024, aes(x = coalition.support, y = sum_apartments)) +
  geom_point(alpha = 0.2, color = "#740410") +  # Increase alpha transparency and change point color
  geom_smooth(aes(color = "Smoothing Curve"), 
              method = 'glm', method.args = list(family = 'poisson'), se = TRUE, 
              linetype = "solid", size = 1.5) +  # Adjust line type for smooth curve
  labs(x = "Coalition Support", y = "Sum of Apartments") +
  scale_color_manual(values = c("Smoothing Curve" = "#740410"), guide = FALSE) +  # Manually set color for smooth curve and hide legend
  theme_minimal() +  # Use minimal theme for cleaner appearance
  theme(plot.title = element_text(size = 16, face = "bold"),  # Adjust title size and style
        axis.title = element_text(size = 14),  # Adjust axis label size
        axis.text = element_text(size = 12)) +  # Adjust axis tick label size
  annotate("text", x = Inf, y = Inf, label = paste("RMSE:", round(rmse, 2)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black")  # Annotate RMSE on the plot
```

