---
title: "Project proposal"
author: "Team number 4"
mainfont: Roboto
fontsize: 10pt
output:
  pdf_document:
    includes:
      in_header: header.tex
# output: html_document
---


```{r load-packages, message=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(htmltools)
library(janitor)
library(kableExtra)
```

```{r setup, include = FALSE}
opts_chunk$set(echo=FALSE) # hide source code in the document
knitr::opts_chunk$set(fig.width=7, fig.height=3.5)
```

## 1. Introduction

Public housing is government-owned housing units provided to low-income individuals and families who cannot afford market-rate housing. Currently, public housing companies which are mostly government-funded, own over 50,000 apartments across Israel. The Ministry of Construction only allows eligible individuals to obtain housing in their registered city, leading to around 4,300 people on the waiting list. Additionally, 25,000 more individuals, including 6,000 Holocaust survivors, face an average wait of 3 to 4 years, with longer waits in Jerusalem and central regions. This situation prompts questions about the government's public housing policy, which our project aims to address:  
**Is there a relation between general elections results per city, to the public housing procurement policy in the same city?**  
Our project aims to find a relation (if exist) between the elected government policy regarding public housing, to the top voted parties in each city.  
Does the government buy more apartments in cities who voted for her? Does it buy less apartments in cities which didn’t vote for her?
In which cities does the government sell more housing units?
In which regions there are more vacant apartments? Did those regions vote for the elected government?
Since 2015, Israel has had 5 national elections, 4 different governments, and 7 different Construction Ministers. While previous research focused on apartment shortages and long waiting times, we aim to investigate if there's a hidden government policy influenced by the vote rate in each city. We will combine public housing data (property purchases, vacant apartments, etc.) and election data (most elected parties per city) to find patterns indicating such policies.

```{r include=FALSE}
Election_results <- read.csv("../data/Translated/elect_results_t.csv")

Properties_rooms <- read.csv("../data/Translated/properties_by_rooms_t.csv")

Purcheuses <- read.csv("../data/Translated/property_purchases_t.csv")

Vacant_apartments <- read.csv("../data/Translated/vacant_apartments_t.csv")

Election_results

Properties_rooms

Purcheuses

Vacant_apartments
```

```{r}
# Renaming the Hebrew variables:
Election_results <- Election_results %>% 
  rename(
    Avoda = "אמת",
    Hofesh_Calcali = "אצ",
    Habait_Hayhudi = "ב",
    Yahadut_Hator = "ג",
    Balad = "ד",
    Hadash = "ום",
    Shahar_Koach = "ז",
    Kama = "זך",
    Koach_Lehaspia = "זנ",
    Tsomet = "זץ",
    Tsionot_Datit = "ט",
    Israel_Hofshit = "י",
    Hacalalit = "יז",
    Ihud_Benei_Brit = "ינ",
    Tsav_Hashaa = "יץ",
    Gush_Tanachi = "יק",
    Machane_Mamalachti = "כן",
    Ani_VeAta = "ך",
    Israel_beitenu = "ל",
    Halicud = "מחל",
    Merets = "מרצ",
    Kvod_Haadam = "נז",
    Nativ = "ני",
    Shema = "נף",
    Atsmaiem = "נץ",
    Yesh_kivun = "נק",
    Anachnu = "נר",
    Reshima_Aravit = "עם",
    Yesh_atid = "פה",
    Piratim = "ף",
    Tseirim_boarim = "צ",
    Manhigut_Hevratit = "ץ",
    Kol_Hasviva = "ק",
    Halev_hayehudi = "קי",
    Seder_hadash = "קך",
    Kol_koveaa = "קנ",
    Beomets = "קץ",
    Reshimat_34 = "רז",
    Shas = "שס",
    Ale_yarok = "ת"
  )

# mutate the names of the cities that are in other datasets.
Election_results <- Election_results %>% 
  mutate(
    settlement.name = case_when(
    settlement.name =="horizons" ~ "Ofakim",
    settlement.name =="The view of the Galilee" ~ "Nof Hagalil",
    settlement.name == "my peace" ~  "Shlomi" ,
    settlement.name == "routes" ~ "Netivot",
    TRUE ~ settlement.name  # Keeps other values unchanged
    )
  )
```

```{r}
Properties_rooms_Citynames <- Properties_rooms[!duplicated(Properties_rooms$CityLMSName_Translated),]

#mutate the names of the cities
Properties_rooms <- Properties_rooms %>% 
  mutate(
    CityLMSName_Translated = case_when(
    CityLMSName_Translated =="horizons" ~ "Ofakim",
    CityLMSName_Translated =="The view of the Galilee" ~ "Nof Hagalil",
    CityLMSName_Translated == "my peace" ~  "Shlomi" ,
    CityLMSName_Translated == "routes" ~ "Netivot",
    TRUE ~ CityLMSName_Translated  # Keeps other values unchanged
    )
  )

# replace the columns with translated columns
Properties_rooms$CityLMSName <- Properties_rooms$CityLMSName_Translated
Properties_rooms <- Properties_rooms %>% clean_names()

# remove duplicated translated columns 
Properties_rooms <- Properties_rooms[, !grepl("translated$", names(Properties_rooms))]

# check city names:
# Purcheuses_names <- Purcheuses[!duplicated(Purcheuses$Lamas_name_translated),]
# Purcheuses_names <- Purcheuses_names[c("Lamas_name","Lamas_name_translated")]
# Purcheuses_names

# preform the same for the other datasets
Purcheuses <- Purcheuses %>% 
  mutate(
    Lamas_name_translated = case_when(
    Lamas_name_translated =="horizons" ~ "Ofakim",
    Lamas_name_translated =="The view of the Galilee" ~ "Nof Hagalil",
    Lamas_name_translated == "my peace" ~  "Shlomi" ,
    Lamas_name_translated == "routes" ~ "Netivot",
    Lamas_name_translated == "streets" ~ "Rehovot",
    Lamas_name_translated == "רחובות" ~ "Rehovot",
    Lamas_name_translated == "fresh" ~ "Raanana",
    TRUE ~ Lamas_name_translated  # Keeps other values unchanged
    )
  )

Purcheuses$Lamas_name <- Purcheuses$Lamas_name_translated
Purcheuses <- Purcheuses[, !grepl("translated$", names(Purcheuses))]
Purcheuses <- Purcheuses %>% clean_names()

# Vacant_apartments_names <- Vacant_apartments[!duplicated(Vacant_apartments$CityLmsName_Translated),]
# Vacant_apartments_names <- Vacant_apartments_names[c("CityLmsName","CityLmsName_Translated")]
# Vacant_apartments_names


Vacant_apartments <- Vacant_apartments %>% 
  mutate(
    CityLmsName_Translated = case_when(
    CityLmsName_Translated =="horizons" ~ "Ofakim",
    CityLmsName_Translated =="The view of the Galilee" ~ "Nof Hagalil",
    CityLmsName_Translated == "my peace" ~  "Shlomi" ,
    CityLmsName_Translated == "routes" ~ "Netivot",
    CityLmsName_Translated == "my peace" ~ "Shlomi",
    TRUE ~ CityLmsName_Translated  # Keeps other values unchanged
    )
  )



Vacant_apartments$CityLmsName <- Vacant_apartments$CityLmsName_Translated
Vacant_apartments$CompanyName <- Vacant_apartments$CompanyName_Translated
Vacant_apartments$StatusName <- Vacant_apartments$StatusName_Translated
Vacant_apartments$PropertyTypeName <- Vacant_apartments$PropertyTypeName_Translated
```

```{r}

Vacant_apartments <- Vacant_apartments[, !grepl("translated$", names(Vacant_apartments))]
Vacant_apartments <- Vacant_apartments %>% clean_names()
```

```{r}
#renaming elections table:
Election_results <- Election_results %>% 
  mutate(
    settlement.name = case_when(
    settlement.name =="horizons" ~ "Ofakim",
    settlement.name =="The view of the Galilee" ~ "Nof Hagalil",
    settlement.name == "my peace" ~  "Shlomi" ,
    settlement.name == "routes" ~ "Netivot",
    settlement.name == "my peace" ~ "Shlomi",
    settlement.name == "streets" ~ "Rehovot",
    settlement.name == "רחובות" ~ "Rehovot",
    settlement.name == "fresh" ~ "Raanana",
    TRUE ~ settlement.name  # Keeps other values unchanged
    )
  )


```

```{r include=FALSE, paged.print=TRUE}
Vacant_apartments


Purcheuses <- Purcheuses %>% 
  rename(
    city_lms_name = "lamas_name"
  )
Purcheuses

Properties_rooms
```
 
```{r include=FALSE, paged.print=TRUE}
results_head <- head(Election_results,8)
# knitr::kable(results_head, col.names = gsub("[.]", " ", names(results_head)))
results_head
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
properties_head <- head(Properties_rooms,8)
# knitr::kable(properties_head, col.names = gsub("[.]", " ", names(properties_head)))
properties_head
```
 
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
purcheses_head <- head(Purcheuses,8)
# knitr::kable(purcheses_head, col.names = gsub("[.]", " ", names(purcheses_head)))
purcheses_head
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
vacant_apartments_head <- head(Vacant_apartments,8)
# knitr::kable(vacant_apartments_head, col.names = gsub("[.]", " ", names(vacant_apartments_head)))
vacant_apartments_head
```

```{r include=FALSE}
Vacant_apartments_by_company <- Vacant_apartments %>% group_by(company_name) %>%  mutate(count = n())

Vacant_apartments_by_company <- select(Vacant_apartments_by_company,company_name,count) 
Vacant_apartments_by_company <- distinct(Vacant_apartments_by_company, .keep_all = TRUE)

```

```{r include=FALSE, paged.print=TRUE}
summary(Vacant_apartments)
summary(Properties_rooms)
summary(Purcheuses)
```

```{r include=FALSE}
Properties_rooms
columns_subset <- Properties_rooms[, 3:ncol(Properties_rooms)]
sum_by_room <- colSums(columns_subset)
total_sum <- sum(sum_by_room)
column_percentages <- (sum_by_room / total_sum) * 100
column_percentages
sum_by_room
```

## 3. Preliminary results 
 
Property purchases: 
The **mean cost** of a 3 rooms apartment is 1,405,986 NIS, and the mean cost of a 4 rooms apartments is 1,671,485 NIS.  
Vacant apartments:  
There are **5 companies** that own available apartments, Amidar with 233 apartments, Amigur with 34, Co. Ltd with 8 and flint and shakmona with 1 each. 
Most of the apartment inventory consists of **2, 3 or 4 rooms** apartments, with a distribution of 29.3%, 48.6% and 15% accordingly.  


```{r echo=FALSE}
Vacant_apartments_count <- Vacant_apartments %>% group_by(city_lms_name) %>%  mutate(count = n())


custom_colors <- c("1" = "#4281A4", "1.5" = "#9CAFB7", "2" = "#EAD2AC", "2.5" = "#E6B89C", "3" = "#FE938C", "3.5" = "#DC3F34", "4" = "#870E06", "5" = "#3E3E3B")

ggplot(data = Vacant_apartments_count,
       mapping = aes( x= reorder(city_lms_name ,-count), fill = factor (num_of_rooms))) +
  geom_bar() + #fill='steelblue'
  scale_fill_manual(values = custom_colors) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  labs(x = "City", y = "Aartments", title = "Number of available apartments by city",fill = "Number of rooms",subtitle = "collected in May 24")
```
```{r echo=FALSE}

custom_colors <- c("2017" = "#4281A4", "2018" = "#9CAFB7", "2019" = "#EAD2AC", "2020" = "#E6B89C", "2021" = "#FE938C", "2022" = "#DC3F34", "2023" = "#870E06", "2024" = "#3E3E3B")

ggplot(data = Purcheuses,
       mapping = aes( x= city_lms_name, fill = factor(year))) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0))+
  labs(x = "City", y = "Apartments bought", title = "Apartments bought each year by City", fill = "Year", subtitle = "From 2017 to 2024")
```

```{r echo=FALSE}
Properties_room_sum <- Properties_rooms %>% 
  group_by(city_lms_name) %>% 
  summarise(total = sum(one_rooms + two_rooms + three_rooms + four_rooms + five_rooms + six_rooms + seven_rooms +  eight_rooms + nine_rooms + ten_rooms + more_rooms, na.rm = TRUE))

Properties_room_sum_filtered <- Properties_room_sum %>%
  filter(total > 300)

Properties_room_sum_filtered2 <- Properties_room_sum %>%
  filter(total < 100)


# Create the bar plot with color gradient fill
ggplot(data = Properties_room_sum_filtered, mapping = aes(x = reorder(city_lms_name, -total), y = total, fill = total)) +
  geom_col(width = 0.7, position = position_dodge(width = 0.9)) +
  scale_fill_gradient(low = "grey", high = "red") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5, margin = margin(t = 10))) +
  labs(x = "City name", y = "Total apartments", title = "Total apartments by City", subtitle = "Where there are more than 300 apartments in the city") +
  theme(plot.margin = margin(10, 10, 10, 10),
        axis.line = element_line(colour = "darkgrey")) +
  scale_x_discrete(expand = expansion(mult = c(-1, -1)))

```


Looking at the data visualization and some summary statistics we can see that there are some concentrations for the current public housing system, where a major portion of the apartments are 2 and 3 rooms which may not help those who need theme. Furthermore there is 1 company owning most of the apartments and a lot of the apartments are located in southern Israel in cities such as Be'er Sheva or Dimona.  
The elections results dataset is used mainly to determain the most voted party in a city, thus we didn't include any explanatory statistics about it.   

## 4. Data analysis plan
#### 1. Outcome (Response, Y) and Predictor (Explanatory, X) Variables

a. First Relation to examine : Outcome (Y): The number of houses bought per city per year. Explanatory (X): Segmentation of voting percentages in that city.
b. Second Relation to examine : Outcome (Y): The number of houses bought in a certain city in relation to the percentage of apartments available out of the total number of existing apartments in the city. Explanatory (X): Segmentation of voting percentages in that city.

#### 2.  Comparison Groups  

  * Temporal Comparison: Different years where election results were different.
  * Spatial Comparison: Different cities in the same year with varying voting patterns.

#### 3. Methods to Answer the Question  

* **Data Preprocessing:** Ensuring all data is uniform, Predicting or imputing null values to complete the dataset. Removing duplicates, correcting errors, and ensuring consistency. Creating new variables or modifying existing ones to better fit the analysis
* **Descriptive Analysis:** Describe the current public housing procurement policy, its components, and influencing factors. make variable overview: Detail the apartment inventory, political state of the city, and other relevant variables.
* **Statistical Analysis:**
  + Logistic Regression: To model the probability of a certain number of houses being bought based on voting percentages. 
  + Linear Regression: To measure the linear relationship between the number of houses bought and the voting percentages.
  + Comparative Analysis: Using t-tests or ANOVA to compare housing procurement policies across different cities and years based on their voting patterns.
  + Correlation Analysis: To check the strength and direction of the relationship between voting patterns and housing purchases.


#### 4. Necessary Results to Support the Hypothesis  

a. Comparison Analysis: Identify and quantify differences in public housing procurement policies by comparing cities with different political leanings.
b. Inventory Analysis: Measure the percentage of vacant apartments and correlate this with housing purchases.


#### 5. Teamwork and Division of Work  

* Dori: Data Collection and Preprocessing - Gather and clean data, handle translations, manage missing values.
* Daniel: Descriptive and Exploratory Analysis - Conduct descriptive analysis, visualize data, and provide initial insights.
* Ofek: Statistical Analysis - Perform logistic, linear, and non-linear regression analyses, comparative and correlation analyses.
* Shahaf: Reporting and Presentation - Consolidate findings, prepare the final report, and create visual presentations for clarity and communication of results.  

To assure all team members experience every part of the data science, each task completed will be reviewed and confirmed by all team members before granting "Finished" status and merging to the project's Master version.  
In addition, we present a general division of work, which will change dynamicaly through the project's progress according to the team's demands and needs.


 
#### Collaboration and Communication:
Tedious work will happen offline where each person will work at home. We will share a Git repository to keep the R code updated. Bi-weekly Zoom meetings will be held to go over the tasks needed to be done, ensuring everyone is on track and any issues can be addressed promptly. This regular communication is essential, especially considering Daniel's Miloim (reserve duty) commitments. During these meetings, we will review progress, plan next steps, and ensure alignment among all team members. By maintaining clear and consistent communication, we aim to work efficiently and effectively, despite being in different locations.



## Appendix

### Data README

```{r include_data_readme, comment=''}
cat(readLines('../data/README.md'), sep = '\n')
```


